Mixed effect tweedie model in TMB

### 2019-04-23 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
test if meanSST does any better than SST--yes, but not enough to really be useful

Load data
```{r}
# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
# nrow(d_trawl1)
d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-SST.rds"))
nrow(d_trawl)
# d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
unique(d_trawl$year)
d_trawl[is.na(d_trawl$year), ]$year <- 2003
d_trawl <- d_trawl[!is.na(d_trawl$depth_m), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
d_trawl <- mutate(d_trawl,
  depth_scaled = arm::rescale(log(depth_m)),
  SST_scaled = arm::rescale(SST),
  depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth_m>50,0,1)
)
glimpse(d_trawl)
```

@@

```{r}
ggplot(dplyr::filter(d_trawl, depth_m <50), aes(SST,temperature_c, colour = depth_m)) + geom_point()
```

@@

```{r}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3_nosst <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

@@

predictions3 <- predict(m_temp3)
predictions3b <- predict(m_temp3_nosst)

predictions3$data$residuals <- residuals(m_temp3)
predictions3b$data$residuals <- residuals(m_temp3_nosst)

@@

ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
  geom_point() + geom_abline(intercept = 0, slope = 1)
```

@@

Short example: 
```{r, eval=FALSE}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  time_varying = list(NULL),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(135, 140, 145, 150),
  seed = c(1)
)

@@

## Just compare fixed slope models with simplest random effect structure
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 1 + SST + depth_scaled,
    temperature_c ~ 1 + SST + depth_scaled + depth_scaled2,
    temperature_c ~ 1 + SST + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled,
    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled + SST * depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled + SST * depth_scaled2 + SST * depth_scaled3
  ),
  time_varying = list(NULL),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

@@

## Compare time-varying slope models still with simplest random effect structure
All of top models for any seed contained year and none showed a linear effect of depth
All are better than fixed slope models
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100), # keep same n_knots
  seed = c(1111, 2222, 3333, 4444) # keep same seeds
)

@@

## Now vary random effect configuation for top model(s) 
a separate spatial field is in all models with highest likelihoods for each seed
AR1 doesn't make much difference (differences between years appear independent)
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(FALSE, TRUE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

@@

```{r}
cv_args <- generate_cv_args(
    formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(TRUE),
  n_knots = c(80,120,140),
  seed = c(1111, 2222, 3333, 4444)
)

@@

View(more_knots$summary)
write.csv(more_knots$summary, file = "cv-output/cv-more-seeds.csv")
saveRDS(more_knots, file = "cv-output/cv-more-seeds.rds")
glimpse(more_knots$summary)
```

## Change seeds 
Also increased knot number too 
Model without SST is always very slightly higher ll

```{r}
cv_args <- generate_cv_args(
    formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(TRUE),
  n_knots = c(130),
  seed = c(5555, 6666, 7777, 8888, 9999)
)

@@

Current top model
```{r}
plan(multiprocess)
temp_rw <- sdmTMB_cv(
  data = d_trawl, n_knots = 130, x = "X", y = "Y", k_folds = 10,
  formula = temperature_c ~ 0 + as.factor(year) + SST,
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", family = gaussian(link = "identity"),
  ar1_fields = FALSE, include_spatial = TRUE, seed = 999
)
sum(temp_rw$sum_loglik)
```

@@

Does a fixed interaction with SST do any better?

```{r}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + SST*depth_scaled + SST*depth_scaled2 + SST*depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

>>>>

Load data
```{r}
# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/data/sensor-data-processed.rds"))
# nrow(d_trawl1)
d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST.rds"))
nrow(d_trawl)
# d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
unique(d_trawl$year)
d_trawl <- d_trawl[!is.na(d_trawl$depth_m), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
d_trawl <- mutate(d_trawl,
  depth_scaled = arm::rescale(log(depth_m)),
  SST_scaled = arm::rescale(meanSST),
  depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth_m>40,0,1)
)
d_trawl[is.na(d_trawl$shallow), ]
d_trawl[is.na(d_trawl$SST_scaled), ]

glimpse(d_trawl)
```

@@

```{r}
ggplot(dplyr::filter(d_trawl, depth_m <60), aes(SST,temperature_c, colour = -depth_m)) + geom_point()
```

Explore mean  SST 
```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "meanSST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r}
ggplot(dplyr::filter(d_trawl, depth_m <50), aes(meanSST,temperature_c, colour = -depth_m)) + geom_point()
```

@@

```{r}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + SST_scaled*shallow, 
  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3b <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year), 
  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

@@

predictions3 <- predict(m_temp3)
predictions3b <- predict(m_temp3b)
#predictions3b <- predict(m_temp3_nosst)

predictions3$data$residuals <- residuals(m_temp3)
predictions3b$data$residuals <- residuals(m_temp3b)

@@

ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
  geom_point() + geom_abline(intercept = 0, slope = 1)
```
Residual plots:
```{r}
ggplot(predictions3$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
ggplot(predictions3b$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-7,8) +
  geom_smooth()
ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
  geom_point() +
  ylim(-7,8) +
  geom_smooth()
```

@@

Short example: 
```{r, eval=FALSE}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  time_varying = list(NULL),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(135, 140, 145, 150),
  seed = c(1)
)

@@

## Just compare fixed slope models with simplest random effect structure
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled,
    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 1 + SST_scaled + depth_scaled,
    temperature_c ~ 1 + SST_scaled + depth_scaled + depth_scaled2,
    temperature_c ~ 1 + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled,
    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled + SST_scaled * depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled + SST_scaled * depth_scaled2 + SST_scaled * depth_scaled3
  ),
  time_varying = list(NULL),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

@@

## Compare time-varying slope models still with simplest random effect structure
All of top models for any seed contained year and none showed a linear effect of depth
All are better than fixed slope models
```{r}
cv_args <- generate_cv_args(
  formula = list(
    #temperature_c ~ 0 + SST_scaled*shallow,
    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
    #temperature_c ~ 1 + SST_scaled,
    temperature_c ~ 0 + as.factor(year) + SST_scaled,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100), # keep same n_knots
  seed = c(1111, 2222, 3333, 4444) # keep same seeds
)

@@

## Now vary random effect configuation for top model(s) 
a separate spatial field is in all models with highest likelihoods for each seed
AR1 doesn't make much difference (differences between years appear independent)
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
    temperature_c ~ 0 + as.factor(year) + SST_scaled,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(FALSE, TRUE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

@@

```{r}
cv_args <- generate_cv_args(
    formula = list(
    # temperature_c ~ 0 + as.factor(year) + SST_scaled,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE,TRUE),
  include_spatial = c(TRUE),
  n_knots = c(140,180,220,250),
  #seed = c(1111, 2222, 3333, 4444) 
  seed = c(5555, 6666, 7777, 8888, 9999)
)

@@

View(more_knots$summary)
write.csv(more_knots$summary, file = "cv-output/cv-more-knots2.csv")
saveRDS(more_knots, file = "cv-output/cv-more-knots2.rds")
glimpse(more_knots$summary)
```

## Change seeds 
Also increased knot number too 
Models without SST_scaled are usually better

```{r}
cv_args <- generate_cv_args(
    formula = list(
    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
    temperature_c ~ 0 + as.factor(year) + SST_scaled,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(TRUE, FALSE),
  n_knots = c(130),
  seed = c(5555, 6666, 7777, 8888, 9999)
)

@@

Current top model
```{r}
plan(multiprocess)
temp_rw <- sdmTMB_cv(
  data = d_trawl, n_knots = 130, x = "X", y = "Y", k_folds = 10,
  formula = temperature_c ~ 0 + as.factor(year) + SST_scaled,
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", family = gaussian(link = "identity"),
  ar1_fields = FALSE, include_spatial = TRUE, seed = 999
)
sum(temp_rw$sum_loglik)
```

```{r}
spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 220)
plot_spde(spde)
```

@@

Does a fixed interaction with SST_scaled do any better?

```{r}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + SST_scaled*depth_scaled + SST_scaled*depth_scaled2 + SST_scaled*depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)
